---
title: "HW9"
author: "Arturo Casillas"
date: "October 30, 2017"
output: 
  html_document:
  #  theme: 'readable'
    keep_md: true
---

## Intro

Problems 1 and 2 are here. The full tables after cleaning can be found in the appendix. The link for the files on github is:


https://github.com/ArturoCasillas/Acasillas-SMU-MSDS-homework


#1. Harry Potter Cast

###a.-b. In the IMDB, there are listings of full cast members for movies. Navigate to http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1. Feel free to View Source to get a good idea of what the page looks like in code. Scrape the page with any R package that makes things easy for you. Of particular interest is the table of the Cast in order of crediting.

I initalized the packages and saved the data into a data.frame object below. I only did very basic cleaning

```{r no1_Init, echo=TRUE, message=FALSE}
#1.a. I clicked on the links to view the data
#     The cleaning will be 
#install.packages("XML")
library(XML)
library(xml2)
library(RCurl)
library(rvest)
library(knitr)

#1.b. I will install the packages and import/scrape the data in a very basic way
#      followed the Demo
URL1<-getURL('http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1')
page1<-read_html(URL1)
table1<-html_table(html_nodes(page1, 'table'))

#    Check out what I have to look at
#str(table1)
#    It appears the third 'node' is what I am after
#    I know this from 1.a. when I looked at the htmls
test<-data.frame(table1[3])
str(test)

```

###c.Clean up the table

#### -It should not have blank observations or rows, a row that should be column names, or just '.'

#### -It should have intuitive column names (ideally 2 to start - Actor and Character)

#### -In the film, Mr. Warwick plays two characters, which makes his row look a little weird. Please replace his character column with just "Griphook / Professor Filius Flitwick" to make it look better. . One row might result in "Rest of cast listed alphabetically" - remove this observation.

Now that it is loaded, I save the Cast data set with minor cleaning

```{r no1C_clean, echo=TRUE}

#1.c.
Cast<-data.frame(test[2:length(test[,2]), c(2,4)])
names(Cast)<-c('Actor','Character')
#str(Cast)
#kable(head(Cast))
head(Cast)
tail(Cast)
```

I need to do more cleaning on the data.

```{r no1C_clean2, echo=TRUE}
#Clean the second column
#tail(Cast)
Cast$Character<-gsub('\n  ','',Cast$Character)
Cast=Cast[!grepl('alphabetically:', Cast$Character),]

#DIm, head and tail
#dim(Cast)
head(Cast)
tail(Cast)
```

After basic cleaning, I fix Warick Davis' row

```{r no1C_warwick, echo=TRUE}
###Fix Warick Davis
#find row
warwick=Cast[grepl('/ ', Cast$Character),]
#Create index to see rows I changed
test.indexes<-c((grep('/', Cast$Character)-1):(grep('/', Cast$Character)+2))

#Fix
warwick=cbind.data.frame(  warwick[,1], trimws(strsplit(warwick[,2], '/')[[1]]))
names(warwick)=names(Cast)
#warwick

## Merge back
Cast=rbind.data.frame(Cast[1:(grep('/', Cast$Character)-1),],                       #Before warwick
                      warwick,                                                      #New warwick
                      Cast[(grep('/', Cast$Character)+1):length(Cast$Character),])  #After Warwick
row.names(Cast)<-1:length(Cast$Actor)

#Look at the rows I changed
testR = Cast[test.indexes,]
testR
```


###d. Split the Actor's name into two columns: FirstName and Surname. Keep in mind that some actors/actresses have middle names as well. Please make sure that the middle 2 names are in the FirstName column, in addition to the first name (example: given the actor Frank Jeffrey Stevenson, the FirstName column would say "Frank Jeffrey.")

I will create a function to do this. 

```{r no1D_test, echo=TRUE}
#1.d.
###Create new first and last name columns
#create first and last name function
#THis is to get the data in the form of a vector

first.last <- function(x, sep=' '){
  
#x<-Cast$actor[80:86]
D<-cbind(
         rep( NA_character_, length(x)), 
         rep( NA_character_, length(x))
         )

for(i in 1:length(x)){
  
  #i = 2
  y<-strsplit(x[i], sep)[[1]]                           #split name
  first<-paste(y[-length(y)], collapse=' ')             #get fitst name (not last name)
  last<-y[length(y)]                                    #get last name
  
  D[i,]<-cbind(first, last)                             #combine
                      }

#format and output
  out=data.frame(D)
  names(out)<-c('FirstName','SurName')
  out
}

```


After creating a function to separate into first and last names, I will implement it on the cast data.

```{r no1D_implement, echo=TRUE}
#1.d. Continued
testF<-cbind.data.frame(first.last(Cast$Actor), Cast$Character)
#str(testF)
Cast<-testF
names(Cast)<-c('FirstName','SurName', 'Character')
str(Cast)
```

####e. Present the first 10 rows of the data.frame() - It should have only FirstName, Surname, and Character columns.

The first ten columns are below. The full table is also in the appendix.

```{r no1E, echo=TRUE}

head(Cast)

```


# 2.

###a. On the ESPN website, there are statistics of each NBA player. Navigate to the San Antonio Spurs current statistics (likely http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs). You are interested in the Shooting Statistics table.

###b. Scrape the page with any R package that makes things easy for you. There are a few tables on the page, so make sure you are targeting specifically the Shooting Statistics table.

I will scrape the table like in 1a.-1b.

```{r no2AB, echo=TRUE}

#2.a.-b.
#     Like the Demo
URL2<-getURL('http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs')
page2<-read_html(URL2)
table2<-html_table(html_nodes(page2, 'table'))

#    Check out what I have to look at
#str(table2)
#    It appears the third 'node' is what I am after
#    I know this from 2.a. when I looked at the htmls
test2<-data.frame(table2[2])
str(test2)

```


###c. Clean up the table (You might get some warnings if you're working with tibbles)

#### -You'll want to create an R data.frame() with one observation for each player. Make sure that you do not accidentally include blank rows, a row of column names, or the Totals row in the table as observations.

#### -The column PLAYER has two variables of interest in it: the player's name and their position, denoted by 1-2 letters after their name. Split the cells into two columns, one with Name and the other Position.

#### -Check the data type of all columns. Convert relevant columns to numeric. Check the data type of all columns again to confirm that they have changed!


```{r no2C_basic, echo=TRUE}
#2.c.
##Get Names and remove nonsense first rows
names(test2)=test2[2,]
shootStats<-data.frame(test2[3:(length(test2[,2]) - 1), ])

#Better Names
names(shootStats)<-   #'Player'
  c('Player', 'FG.Made', 'FG.Attempt', 'FG.Percent', 'FG3P.Made', 
    'FG3P.Attempt','FG3P.Percent', 'FT.Made', 'FT.Attmempt',
    'FT.Percent', 'FG2P.Made', 'FG2P.Attempt','FG2P.Percent',
    'Points.per.Sot', 'Adj.FG.Percent');
row.names(shootStats)<-1:13
```

Now to separate player and position in the first column

```{r no2C_sep, echo=TRUE}
#Separate player and position
#I will recycle my earlier funciton but I have to change the names
testP<-first.last(shootStats[,1], sep=',')
names(testP)=c('Name', 'Position')
shootStats=cbind.data.frame(testP, shootStats[,-1])
```

Now to convert character columns to numeric

```{r no2C_num, echo=TRUE}
###Convert character to numeric
#CHeck shootStats
str(shootStats)

#use apply() function
m<-apply(shootStats[,-c(1:2)], 2, as.numeric)

#merge back
shootStats=cbind.data.frame(shootStats[,c(1:2)], m)

#Check shootStats again
str(shootStats)
head(shootStats)

```

###d. Create a colorful bar chart that shows the Field Goals Percentage Per Game for each person. It will be graded on the following criteria.

#### -Informative Title, centered

#### -Relevant x and y axis labels (not simply variables names!)

#### -Human-readable axes with no overlap (you might have to flip x and y to fix that). Note: You do not have to convert the decimal to a percentage.

#### -Color the columns by the team member's position (so, all PF's should have the same color, etc.)

I can use ggplot to accomplish this

```{r, no2e}
# BarPlot #
library(ggplot2)

ggplot(dat=shootStats, aes(x=Name, y=FG2P.Percent, fill=Position)) + 
  geom_bar(stat='identity') + coord_flip()  + theme_minimal() +                             #Format
  ggtitle("Plot of Field Goal Percentage \n per Player per Game") +                         #Title
  theme(plot.title = element_text(hjust = 0.5)) +                                           #Center Title
  theme(legend.position = c(0.9, 0.5)) +                                                    #Legend Position
  xlab("Player Name") + ylab("Field Goal % per Game") +                                     #axis labels
  scale_fill_manual(values=c("#000000", "#005511", "#000099", "#777777", "#551100")) #+     #Manual set colors
  #scale_color_gradient(low = "#00000000", high = "#0091ff")
```

This is a link to my GitHub repository where the files and codebook are contained:
https://github.com/ArturoCasillas/Acasillas-SMU-MSDS-homework

#Appendix

Full Cast Data
```{r}
kable(Cast)
```

Full Shooting Statistics Data
```{r}
kable(shootStats)
```