---
title: "HW4"
author: "Arturo Casillas"
date: "September 26, 2017"
output: 
  html_document:
    keep_md: true
---

## Intro

Problems 1.-3. are here. The swirl document is in a separte file. 

#1. 

a. Install the fivethirtyeight package.

```{r no1A, echo=TRUE}
#2.a.
#First I install the packages
#install.packages("fivethirtyeight")
#install.packages("devtools")
# I already installed the package so I am including 
library("fivethirtyeight")
```

b. & c. In the listing of Data sets in package 'fivethirtyeight,' assign the eighteenth data set to an object 'df.' 


```{r no1B, echo=TRUE}
#1.b.
#Also, I received the code below from the README
# To see a list of all data sets:
#data(package = "fivethirtyeight")

# To see a more detailed list of all data sets, see the package vignette:
#vignette("fivethirtyeight", package = "fivethirtyeight")

#I will try to list them here
ls("package:fivethirtyeight")


#1.c.
#The 18th data set on the vignette is "college_recent_grads"
ls("package:fivethirtyeight")[18]
#given what no. 2 says below, I'm going with recent grads 
df0<-comma_survey
df<-college_recent_grads
class(df)
```

c. Use a more detailed list of the data sets to write out the URL in a comment to the related news story.

```{r no1C, echo=TRUE}
# I am interpreting this as the URL on the vignette
URL<-"http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/"
URL
```

d. Using R command(s), give the dimensions and column names of this data frame.

```{r no1D, echo=TRUE}
#Here the column names and dimension of the data set
dim(df)
names(df)

#BTW, here are the names and dimensions of the other dataset
dim(df0)
names(df0)
```

# 2.

a. write an R command that gives you the column names of the data frame. Right after that, write one that counts the number of columns but not rows.

```{r no2A, echo=TRUE}
#############################
## 2
#############################

#2.a.#
#column names#
names(df)


#number of columns#
length(names(df))

#alternate attempt
length(df[1,])

```


b. Generate a count of each unique major_category in the data frame. 

```{r no2B, echo=TRUE}
#2.b.#
#assign count of major_category to an object
major_count <- table(df$major_category)
major_count
```

c. Make a barplot of major_count. Make sure to label the title
with something informative, label the x and y axis, and
make it any color other than grey. 

```{r no2C, echo=TRUE}
#2.c.#
#plot the frequency of major_category
par(las=2)
barplot(major_count, main="Number of Grads. Surveyed by Category of Major", horiz=TRUE, xlab="Number Surveyed", ylab="Major Category", col = 4)
#I couldn't figure out how to make the yaxis label show.
```

```{r no2D, echo=TRUE}
#2.d.
write.csv(df, "college_survey.csv", row.names = FALSE)
```

#3.

This is a link to my GitHub repository:
https://github.com/ArturoCasillas/Acasillas-SMU-MSDS-homework.git

It has one branch 

