---
title: "HW5"
author: "Arturo Casillas"
date: "October 4, 2017"
output: 
  html_document:
    keep_md: true
---

## Intro

Problems 1.-3. are here. The link for 4. and the files on github is:


https://github.com/ArturoCasillas/Acasillas-SMU-MSDS-homework


#1. 

a. First, import the yob2016.txt file into R so you can process it. Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.

```{r no1A, echo=TRUE}
#1.a.
library(reshape2)
getwd()
df=read.table("C:/Users/acasi/Downloads/yob2016.txt", sep=";", header=FALSE)
head(df)
#Will add year for later
names(df)=c("Name", "Sex", "No.of.Children.2016")

```


b. Display the summary and structure of df 

```{r no1B, echo=TRUE}
#1.b.

#1.b.
str(df)
summary(df)


```

c. Your client tells you that there is a problem with the raw file. One name was entered twice and misspelled. The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name. Write an R command to figure out which name it is and display it.

```{r no1C, echo=TRUE}
#1.c.
#some backround
max(duplicated(df[,1])*1)

##Here is the command to display the dublicated name
df[grep("yyy", df[,1] ),1]

```

d. Upon finding the misspelled name, please remove this particular observation. Save the remaining dataset as an object: y2016

```{r no1D, echo=TRUE}
#1.d.
y2016=df[-212,]
```



# 2.

a. Like 1a, please import the yob2015.txt file into R. Look at the file before you do. Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.

```{r no2A, echo=TRUE}

#2.a.
df=read.table("C:/Users/acasi/Downloads/yob2015.txt", sep=",", header=FALSE)
head(df)
#Will add year for later
names(df)=c("Name", "Sex", "No.of.Children.2015")
str(df)

```


b. Display the last ten rows in the dataframe. Describe something you find interesting about these 10 rows. 

```{r no2B, echo=TRUE}
#2.b.
tail(df, 10)

##I find it strange that there are exactly 5 children with these names in each case

```

c. Merge y2016 and y2015 by your Name column; assign it to final. The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.

```{r no2C, echo=TRUE}
#2.c.
final=merge(df, y2016, by=c("Name","Sex"))
str(final)
#inner join also meets the both 2016 and 2015 criteria
#Also, 2ds discussion plus the demands of number three suggests I merge by "Name" and "Sex", unlike what it says on the problem text.

#check for missing values
#max(is.na(final[,3]))
#max(is.na(final[,4]))
m=apply(final, c(1,2), is.na)
max(m)

```
# 3.

a. Create a new column called "Total" in final that adds the amount of children in 2015 and 2016 together. In those two years combined, how many people were given popular names?

```{r no3A, echo=TRUE}
#3.a.
final$Total=final$No.of.Children.2015 + final$No.of.Children.2016

#How many were given popular names?
#summary(final$Total)
#I'm going to assume this means how many children had names whose combined total count was above the mean
sum(final$Total[final$Total>mean(final$Total)])

```


b. Sort the data by Total. What are the top 10 most popular names? 

```{r no3B, echo=TRUE}

#3.b.
test=final[order(final$Total, decreasing=TRUE), ]
head(test, 10)
```

c. The client is expecting a girl! Omit boys and give the top 10 most popular girl's names. Assign this to object girl.

```{r no3C, echo=TRUE}
#3.c.
#part 1
girl=head(test[(test$Sex == "F") , ], 10)
#Note any problems with unisex names were resolved in the inner merge above on both "Name" and "Sex"
#Also, I am taking "this object" to mean the 10 most popular girl names
girl
```


d. Write these top 10 girl Names and their Totals to a CSV file called itsagirl.csv. Do not include row labels. Leave out columns other than Name and Total. 

```{r no3D, echo=TRUE}
#3.d.
write.csv(girl[, c(1,5)], "itsagirl.csv", row.names = FALSE)

#For precaution given the instructions
write.csv(test[(test$Sex == "F"), c(1,5)], "itsagirl_all.csv", row.names = FALSE)
```


#4.

This is a link to my GitHub repository where the files and codebook are contained:
https://github.com/ArturoCasillas/Acasillas-SMU-MSDS-homework


